---
title: "Demo script to run clei2block"
output: html_notebook
---

Please make sure the system is equipped with Nvidia GPU and has docker installed. 

Assume this R Notebook is located at clei2block/code and there is the demo-data directory at clei2block/. 

### Train clei2block model
With Nvidia Tesla T4 GPU, training the model takes about 3 minutes.

clei2block.py \<path to input data\> \<path to output\> \<dimensionality of the latent space\> \<dimensionality of hidden layer\> \<if you want to apply the model to other data specify the path to the data (Optional)\> \<input features for latent embedding\> \<input features for linear regression\>
```{bash}
time docker run --rm --gpus all -v $(pwd)/../:/var  pytorch/pytorch:1.2-cuda10.0-cudnn7-runtime  /bin/bash -c "pip install pandas; python /var/functions/clei2block.py /var/demo-data/ /var/demo-out/ 100 800 none mRNA mRNA"
```
#### Install required R libraries
```{r}
install.packages("ggplot2")
install.packages("dplyr")
install.packages("tidyr")
install.packages("reticulate")
library(reticulate)
np=import("numpy")
```
#### Check if the loss is decreased as expected.
```{r}
#install.packages("ggplot2")
#install.packages("dplyr")
#install.packages("tidyr")
library(ggplot2)
library(dplyr)
library(tidyr)

# collect training loss
list.files("../demo-out/",pattern = "train_loss.txt",full.names = TRUE,recursive = TRUE)%>%
  lapply(., function(x){
    read.delim(x,sep = ",",header = F)%>%
      mutate(trial=dirname(x)%>%basename(),
             step=1:n())%>%
      dplyr::select(trial,step,train_loss = V1)%>%
      return()
  })%>%
  bind_rows() -> loss_ds

# collect validation loss
list.files("../demo-out/",pattern = "valid_loss.txt",full.names = TRUE,recursive = TRUE)%>%
  lapply(., function(x){
    read.delim(x,sep = ",",header = F)%>%
      mutate(trial=dirname(x)%>%basename(),
             step=1:n())%>%
      dplyr::select(trial,step,valid_loss = V1)%>%
      return()
  })%>%
  bind_rows()%>%
  inner_join(loss_ds,.,by=c("trial","step"))%>%
  group_by(trial)%>%
  filter(1:n() <= which(valid_loss==min(valid_loss)))-> loss_ds

loss_ds%>%gather(loss_type,loss,-trial,-step)%>%
    ggplot(.,aes(as.factor(step),loss,color=loss_type))+
  geom_boxplot()+
  xlab("Training epochs")
```

#### Check the model performance for testing data
```{r,fig.width=5}
#install.packages("reticulate")
#np=import("numpy")
library(reticulate)
np=import("numpy")

# collect test prediction
list.files("../demo-out/",pattern = "test_prediction.npy",full.names = TRUE,recursive = TRUE)%>%
  lapply(., function(x){
    np$load(x)%>%
      as.matrix()%>%
      t()%>%
      return()
  }) -> pred_protein

# consensus prediction
for(i in 2:length(pred_protein)){
  pred_protein[1][[1]]=pred_protein[1][[1]]+pred_protein[i][[1]]
}
pred_protein = pred_protein[1][[1]] # this is a protein by sample matrix predicted by the clei2block model


# read actual measurements
acutal_protein = read.delim("../demo-data/test_y.txt.gz",header = F)%>%as.matrix()
acutal_mRNA = read.delim("../demo-data/test_lr_x_mRNA.txt.gz",header = F)%>%as.matrix()

# calculate correlation between predicted protein and acutal protein measurement for each gene across samples
lapply(1:nrow(acutal_protein), function(i){
  cor(pred_protein[i,],acutal_protein[i,])
})%>%unlist() -> pre_cor

# calculate correlation between acutal protein measurement and thier corresponding mRNA for each gene across samples
lapply(1:nrow(acutal_protein), function(i){
  cor(acutal_mRNA[i,],acutal_protein[i,])
})%>%unlist() -> mRNA_cor

# visualize correlation
ds = data.frame(cor=pre_cor,
           data = "predicted protein")
ds = bind_rows(data.frame(cor=mRNA_cor,
           data = "actual mRNA"),ds)
ds%>%
  ggplot(.,aes(data,cor,fill=data))+
  geom_boxplot()+ylab("Correlation with actual protein")
```

### Use a pretrained model to predict test data
```{bash}
for ITER in 0 1 2 3 4
do
  time docker run --rm --gpus all -v $(pwd)/../:/var  pytorch/pytorch:1.2-cuda10.0-cudnn7-runtime  /bin/bash -c "pip install pandas; python /var/functions/clei2block-predict.py /var/demo-data/test_ /var/demo-out-predict/$ITER/ /var/demo-out/$ITER/trained_model.pt mRNA mRNA"
done
```


#### Check the model performance for testing data predicted by the pretrained model
```{r,fig.width=5}
#install.packages("reticulate")
#np=import("numpy")
library(reticulate)
np=import("numpy")

# collect test prediction
list.files("../demo-out-predict/",pattern = "prediction.npy",full.names = TRUE,recursive = TRUE)%>%
  lapply(., function(x){
    np$load(x)%>%
      as.matrix()%>%
      t()%>%
      return()
  }) -> pred_protein

# consensus prediction
for(i in 2:length(pred_protein)){
  pred_protein[1][[1]]=pred_protein[1][[1]]+pred_protein[i][[1]]
}
pred_protein = pred_protein[1][[1]] # this is a protein by sample matrix predicted by the clei2block model


# read actual measurements
acutal_protein = read.delim("../demo-data/test_y.txt.gz",header = F)%>%as.matrix()
acutal_mRNA = read.delim("../demo-data/test_lr_x_mRNA.txt.gz",header = F)%>%as.matrix()

# calculate correlation between predicted protein and acutal protein measurement for each gene across samples
lapply(1:nrow(acutal_protein), function(i){
  cor(pred_protein[i,],acutal_protein[i,])
})%>%unlist() -> pre_cor

# calculate correlation between acutal protein measurement and thier corresponding mRNA for each gene across samples
lapply(1:nrow(acutal_protein), function(i){
  cor(acutal_mRNA[i,],acutal_protein[i,])
})%>%unlist() -> mRNA_cor

# visualize correlation
ds = data.frame(cor=pre_cor,
           data = "predicted protein")
ds = bind_rows(data.frame(cor=mRNA_cor,
           data = "actual mRNA"),ds)
ds%>%
  ggplot(.,aes(data,cor,fill=data))+
  geom_boxplot()+ylab("Correlation with actual protein")
```
